Offline Mode for Corporate Mobile Client Security Architecture

### 2016-08-28 docs/papers/2016JGC/2016JGC.tex
primeiros passos para remocao da parte online o paper

\begin{abstract}
Preventing data leakage on the mobile device is a crucial security problem. Therefore, additional control and protection measures should be taken for the confidential data on the mobile devices that leave the boundaries of the organization. In this scenario, such devices can be used on the offline mode for the convenience and traffic reduce or even for a particular characteristic of the app or of the business. This paper presents a novel approach to the security of the corporate mobile devices, in particular the offline mode. The important part of the presented approach is the conceptualization and the definition of the core methodology to solve the problem of offline mobile security, i.e. the protection of the confidential data in use when the mobile client is not connected to the corporate cloud. The protection of the sensitive data is provided by the combination of the cryptographic methods, such as AES file encryption, ABE authorization based both on user and share attributes, user key secret sharing (SS) based protection between the device and the user as well as MOS-based analytics methods to prevent the malicious user behavior. The proposed security architecture supports the basic mobile device protection principles: minimized traffic load and reduced communication with the cloud; usage of the light-weighted operations and an optimized combination of the security methods.
\end{abstract}

\keywords{ mobile security \and offline mode ABE \and Model Order Selection (MOS) \and user behavior analysis \and secret sharing \and protected cloud }

@@

Additional security issues and requirements have to be considered when mobile devices are actively used in corporate cloud environment \cite{yovel2014}. Today more and more organizations and enterprises are functioning in the Bring-Your-Own-Device (BYOD) paradigm. The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS). While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, such as CASB and CAC, the corporate mobile app is usually light-weighted and generally less protected. The protection scheme used on a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}. Therefore, encrypting files and generating keys on a mobile device is not considered a good solution. On the other hand, the protection schemes with good computational qualities lack the security analysis in many cases \cite{khan2014bss}. The common practice is the shadow user activities monitoring \cite{yovel2014}. However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode.

Suppose that a SME uses CASB in order to protect data at rest, i.e. while stored on the server-side, in transit, communicating with server, and, in use, while the client is connected to the network. In this case, when the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile app has to secure itself with its own limited resources. Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection. For example, offline mode does not allow performing the extensive computation and encryption on the mobile device.
Observing the above described landscape, this paper outlines the concept of the offline mobile client security. This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing \cite{galibus2007} and ABE encryption \cite{goyal2006attribute}. Moreover, our proposed approach proposal includes the usage of the user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to highlight possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack. The key expiration period is safely incorporated into the proposed system solution in order to enhance security. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors. The main target of our proposed solution is to provide a maximum defense at the minimal resource cost.

The paper is structured as follows. Section 2 analyzes the most common security problems in the mobile cloud environment and the proposed solutions for them related to the offline protection in the BYOD world. Section 3 outlines the basic processes and the module security infrastructure of the proposed solution as well as the outline of online mode security. Section 4 presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section 5 provides the schemes of the main security blocks including encryption and key usage hierarchies as well as explanations of the core security methods: SSS, ABE and MOS. Section 6 discusses the security proofs including the outline of adversary model, common threat scenarios and the proposed analytics to discover abnormal usage. Still in Section 6, complexity analysis and practical implementation results are included. Section 7 concludes the paper.

\section{Related works}
\label{sec_related_works}
The increasing usage of BYOD demands more sophisticated data protection services compared to ordinary computing environment. A common practice is to provide additional contextual methods apart from authentication, DLP services, and encryption, which can be at rest, in transit and in use \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. The contextual methods increase the security of the app at a maximum level with minimum resource requirements. The most commonly used are:
Using geolocation of the device to trace its usage;
Setting up the expiration period of an app;
Secured transfer politics between apps;
Restricting access to the corporate app;
Setting up the expiration period of app pass pin;
Setting up the counter of failed tries;
Restricted or prohibited offline access;
Logging and auditing.
Preventing data leakage on the mobile device is a crucial security problem. Therefore, administrators should take additional control and protection measures for the confidential data on the mobile devices that leave the boundaries of the organization. Generally, the most sensitive and confidential data should not be permitted to be transferred to the mobile device. However, what if the SME need to allow their employees to work on such devices and even use them on the offline mode for the convenience and traffic reduce or even for a particular characteristic of the app or the business itself?
From the theoretical point of view of this problem, there are several surveys, whose common point is the mobile cloud computing as a rapidly developing paradigm that poses many security and complexity problems \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. An analysis of the new models of mobile cloud computing and new ways of using data storage services is presented in \cite{khan2015cloud, khan2013towards}. Commonly, all the models and protection schemes concentrate on the encryption properties and either perform the computations on their own \cite{zhao2010trusted, yang2011provable} or use the cloud provider to off-load the cryptographic operations \cite{itani2010energy, ren2011lightweight}. Obviously, it is natural the mobile device cannot handle all operations securely without the assistance of a cloud provider, due to resources constraint and battery limitation.
The necessity to use schemes that function without putting load on a provider arises when it is desired to make the device less dependent on the cloud provider, i.e. corporate app continues to provide the secure access to the sensitive data without connection to the network. As discussed in \cite{khan2015cloud}, all the currently known schemes of encryption, performing the computation, either use a cloud provider \cite{zhao2010trusted}, a third party trusted agent \cite{yang2011provable} or a combination of both \cite{itani2010energy}. In some cases, they concentrate on computational complexity without taking care of user privacy and security \cite{ren2011lightweight}.
Therefore, according to \cite{khan2015cloud, tenorio2013greatest, khan2013towards}, the state-of-the-art mobile cloud security models do not consider the problem of the offline security mode. In many cases the industrial providers of the mobile security API avoid the problem by completely forbidding the offline access to the protected app, i.e. SAP Mocona, which operates as a secure app wrapping layer \cite{Lelyveld2013}. Due to various constraints such as traffic load, travelling and ease of access, the SME business procedure may require such access. To the best of our knowledge, the offline mode security problem has not yet been deployed, neither in academia nor in the industry \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. Therefore, the main concern of this proposal is the protection of the device and app in offline mode when the functions of data protection cannot be offloaded to a cloud or a trusted party.

\section{The client security architecture}
\label{sec_the_client_security_architecture}
The approach proposed in this work describes and implements the complete lifecycle of the mobile app security infrastructure. The security processes depend on the key expiry period, and are used to access the protected storage. Once the user keys expire, the user is requested to enter his valid credentials, i.e. PIN and password. The client app then sends the credentials to the server for verification. Once the new set of access keys is received, the user can access the protected files in the offline mode, without the access to the server. This means that no further communication with the server is needed until the key expires. The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:1}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=7cm]{figures/coresetoffunctionsandprotocols.eps}
	\caption{The core set of functions and protocols of the mobile app security infrastructure}
	\label{fig:1}
\end{figure}

@@

The cryptographic functions include the decryption (for AES and ABE) key restoration for SSS and EKE token generation procedure. While the AES-ABE hybrid encryption scheme functions both in the online and offline modes, the SSS is used only in the offline mode and the EKE generation and executions is used in the online mode. These operations are described in Subsections \ref{sec_aes}-\ref{sec_secret} and are part of the Figure \ref{fig:1}, related to online and offline mode operations.

The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and more elaborate MOS-inspired statistics analyzer. Functions such as alerting and deleting the expired key belong to this block as well. These functions are described in Subsection \ref{sec_mos}.

The communication with server includes the separate sender and receiver to check the user credentials and receiver new keys which acts based on the EKE protocol. The remaining data like access list and files can be sent via unprotected channels. Subsection \ref{sec_online_mode} briefly discusses the communication with the server, but it is also an intention to make it a subject of our next work.

The storage is divided in two parts – temporary and permanent. While the files and file keys are stored permanently in order to reduce unnecessary traffic and the resource usage on client, the ABE keys and the key storage protection data are temporary. The unprotected files stored in the mobile client app memory are not the subject of the security architecture.

\subsection{Online mode}
\label{sec_online_mode}
This subsection briefly outlines the main processes and functions in the online mode. The connection to the network allows verifying the user credentials with the help of the protected cloud server. The communication with the cloud is secured by means of the encrypted key exchange protocol \cite{hao2010j, hao2008password} and serves to receive the new set of keys, files and send the collected log information. The client app operations in the online mode are as represented in Figure 3.3.

\begin{figure}[h!]
	\includegraphics[width=8.6cm]{figures/onlinemodeoperations.eps}
	\caption{Online Mode Operations}
	\label{fig:3}
\end{figure}

The communication with server is necessary once the key is expired. This allows optimizing the load on the network once the traffic is reduced to the discrete sessions. In such environment it is important to keep track of the time synchronization both on server and on client in parallel in order to avoid malicious user behavior messing up with time attacks.

As depicted in Figure \ref{fig:4}, the operation of the online mode starts from the client app sending a token to server by means of EKE: DEV\_PASS + TIME + PASS + PIN in order to validate his identity and the identity of the user to the server. This value serves as a proof that the device and the user are the ones with which server communicated previously. After performing the verification the server sends the new calculated DEV\_PASS and TIME to the mobile app by means of established EKE session. Afterwards the server sends the new encrypted key set and encrypted files to the client app in the clear.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8.6cm]{figures/onlineauthorizationworkflow.eps}
	\caption{Online Authorization Workflow}
	\label{fig:4}
\end{figure}

This model proposes to use light-weighted EKE for the fast communication with server communication like J-PAKE \cite{hao2010j, hao2008password} or SIS-based EKE \cite{bogos2013primeless}. SIS is a public key encryption that avoids generating the big primes, thus it can be used for the secure key exchange \cite{bogos2013primeless,bellovin1992encrypted}. EKE is performed as soon as the key is expired and the client asks the user to synchronize with the server. Note that the EKE is not required to send the encrypted files or statistics. The current implementation of mobile app uses J-PAKE protocol. We refer to \cite{hao2010j, hao2008password, toorani2014security} for a complete description and analysis of the J-PAKE protocol.

\section{Offline mode: our proposed model and solution}
\label{sec_offline_mode}
This section proposes an open model of mechanisms for the mobile device protection in which the security is supported both in online and offline modes. Currently and to the best of our knowledge, the systems of mobile device protection follow a model where the protected mobile client can operate only when it is connected to the cloud, which is not always convenient for the end-user. The basic principles of the mobile device protection herein proposed are:

\begin{enumerate}
	\item Optimized communication with the cloud when the device does not need to be constantly connected to the server due to the resource constraint and necessity to secure this communication;
	\item Optimized combination of the security mechanisms so that the mobile client does not need to perform complex computation like encryption and key generation due to its resource constraint;
	\item Behavioral analysis of user's operations on mobile client, which can indicate anomalous or automated activities performed by attackers.
\end{enumerate}

The most important security issues in the proposed model arise when the device goes to the offline mode and the user is still allowed to get the access to the protected SME documents. In this case, the server can neither monitor the user activity nor provide the protection methods. The security should be performed at the mobile client. Additionally, the maximum protection should be provided at the minimum resource cost.

In the online mode the mobile device uses the secure communication with the server in order to verify the validity of user’s credentials. On the contrary, the offline protection model should be approached independently. Thus, the proposal is that the authentication/authorization mechanisms in the offline mode should utilize the derived proof of the user identity. The requirements for the proof are as follows:

\begin{enumerate}
	\item The proof is derived from the previous session, in order to verify that the user is still authorized;
	\item The proof should not give access to user password, i.e. it should not be stored on the device;
	\item The proof should be temporary and have an expiration period;
	\item It should not be directly used in communication with the server, in order to prevent the malicious user from mimicking;
	\item It should be resilient to offline dictionary attacks;
	\item It has to stay effective both in the scenarios of the malicious outsider and leakage of information when the formerly authorized user leaves the group.
\end{enumerate}

Such proof cannot be stored on the client device, as it is not possible to guarantee its protection in the offline mode. Therefore, the most effective way to secure the proof on the offline mode without performing complex computation is to share this proof between the client and the user in a protected manner. In this case, there is no need to store the function of the user password. The additional argument against the traditional password verification is the necessity to check the PIN, which is very small, so the construction of valid one-way function resilient to the offline dictionary attacks is a difficult task.

The proposed offline mode architecture is represented in Figure \ref{fig:5}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8.6cm]{figures/offlinesecurityarchitecture.eps}
	\caption{Online Authorization Workflow}
	\label{fig:5}
\end{figure}

@@

\begin{enumerate}
	\item \textbf{The protected storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files. The outline of the hybrid encryption scheme is presented in the Section \ref{sec_algorithms}.
	\item \textbf{Threat Intelligence Manager (TIM)}: most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode. Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios. Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior according to the scenarios in Section \ref{sec_results}. Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones.
	\item \textbf{Key Management center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
\end{enumerate}

@@

\section{The algorithms, key usage and data protection methods}
\label{sec_algorithms}
In the proposed approach, the kernel encryption scheme in the mobile device is a combination of several methods of security. The files are encrypted with 128/256-bit AES, while the permanent file keys are encrypted with the attribute-based encryption. The set of expiring ABE keys corresponding to the set of files accessible by user in encrypted with a single expiring AES key (KEY\_SET\_KEY). This key is expiring and is splited by the server into 4 parts (2 are stored on the device and 2 belong to the user) by the method of secure secret sharing. The encryption workflow is outlined in the Figure \ref{fig:9}.

@@

The proposed authentication system is based on the shared storing of the user key. Also, the device acts as a dealer in the SSS. Using the SSS ensures that the key can only be accessed by an authenticated user. The participants of the (2, 2)-threshold SSS are the user and device. The user share $s_1(x)$ is computed based on the PIN and the PASS entered by the user. Additionally, the current time value TIME is used in the calculation of the share. Let $s(x) = d$ and $s_1(x) = f(PIN + PASS + TIME)$, where $f$ is a one-way function that transforms the data into the string of the desired length:

@@

The complexity analysis depends on the operation that the user and the mobile app perform in order to keep the device protected. In the offline mode the client app performs the following actions:

@@

The proposed concept of mobile client security has been implemented in the Storgrid protected cloud environment \cite{storgrid2016}. Therefore, the approach is correlated with the practical usability requirements:  the corporate user continues to use the mobile storage app in offline and does not need to reload the files every time the key is renewed. This methodology can be used in other mobile apps. The common advantage is that the mobile client performs the operations both in the offline and online mode and uses the key expiry and ABE to protect the privacy of the corporate data. The Table \ref{tab:01} presents the results of mobile app security workflow testing. Table \ref{tab:01} demonstrates that the device decrypts the data with the acceptable speed. One can compare the speed of ABE/AES decryption (keys and actual data) and the speed of the whole decryption process.

@@

Table \ref{tab:02} demonstrates the results of testing the speed of the mobile device for receiving and displaying the list of files. The app was tested for displaying up to 1000 files. Around 100 requests were generated from 10 threads.

The device receives and displays the list of files depending on the user attributes with acceptable average speed and the proposed protection scheme is usable. Table \ref{tab:03} demonstrates the results of performance testing of the keys generation procedure on server:

@@

It is possible to observe that the processing time increases with the window size decreasing, similar to the results for eigenvalue decomposition time. The longest processing time measured is lower than 200 milliseconds, even considering window size of 10 minutes or processing 6 MB of user operation log. This result represents an acceptable processing time for anomaly detection in mobile devices.

\section{Conclusion and future work }
\label{sec_conclusion}
An important security issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile devices. Once a mobile device releases the connection with the corporate cloud, no security measure implemented in the cloud infrastructure assures the protection of sensitive data stored in the mobile device. Aware of this problem and its importance, this work presented a proposal to address the offline mobile security problem combining a different cryptographic methods. Moreover, this proposed approach also prevents malicious user behavior by applying a MOS-based analytic method.

>>>>

\begin{abstract}
Preventing data leakage on the mobile app is a crucial security problem. Therefore, additional control and protection measures should be taken for the confidential data on the mobile devices that leave the boundaries of the organization. In this scenario, such apps can be used on the offline mode for the convenience and traffic reducing or even for a particular characteristic of the app or of the business. This paper presents a novel approach to the security of the corporate mobile apps, in particular the offline mode. The important part of the presented approach is the conceptualization and the definition of the core methodology to solve the problem of offline mobile security, i.e. the protection of the confidential data in use when the mobile client is not connected to the corporate cloud. The protection of the sensitive data is provided by the combination of the cryptographic methods and analytics methods to detect malicious user behavior. The proposed security architecture supports the basic mobile app protection principles: minimized traffic load and reduced communication with the cloud; usage of the light-weighted operations and an optimized combination of the security methods.
\end{abstract}

\keywords{ offline mobile cloud security \and offline ABE \and Model Order Selection (MOS) \and user behavior analysis \and secret sharing}

@@

Additional security issues and requirements have to be considered when mobile apps are actively used in corporate cloud environment \cite{yovel2014}. Today more and more organizations and enterprises are functioning in the Bring-Your-Own-Device (BYOD) paradigm. The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS). While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, such as CASB and CAC, the corporate mobile app is usually light-weighted and generally less protected. The protection scheme used on a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}. Therefore, encrypting files and generating keys on a mobile device is not considered a good solution. On the other hand, the protection schemes with good computational qualities lack the security analysis in many cases \cite{khan2014bss}. The common practice is the shadow user activities monitoring \cite{yovel2014}. However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode.

Suppose that a SME uses CASB in order to protect data at rest, i.e. while stored on the server-side, in transit, communicating with server, and, in use, while the client is connected to the network. In this case, when the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile app has to secure itself with its own limited resources. Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection. For example, offline mode does not allow performing the extensive computation and encryption on the mobile device.
Observing the above described landscape, this paper outlines the concept of the offline mobile cloud security. This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing (SSS) \cite{galibus2007} and ABE encryption \cite{goyal2006attribute}. Moreover, our proposed approach includes the user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack. The key expiration period is safely incorporated into the proposed system solution in order to enhance security. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors. The main target of our proposed solution is to provide a maximum defense at the minimal resource cost.

The paper is structured as follows. Section \ref{sec_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions related to the offline protection in the BYOD world. Section \ref{sec_the_client_security_architecture} outlines the basic processes and the module security infrastructure of the proposed solution. Section \ref{sec_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section \ref{sec_algorithms} provides the schemes of the main security blocks including encryption and key usage hierarchies as well as explanations of the core security methods: SSS, ABE and MOS. Section 6 discusses the security proofs including the outline of adversary model, common threat scenarios and the proposed analytics to discover abnormal usage. Still in Section 6, complexity analysis and practical implementation results are included. Section 7 concludes the paper.

\section{Related works}
\label{sec_related_works}
The increasing usage of BYOD demands more sophisticated data protection services compared to ordinary computing environment. A common practice is to provide additional contextual methods apart from authentication, DLP services, and encryption, which can be at rest, in transit and in use \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. The contextual methods increase the security of the app at a maximum level with minimum resource requirements. The most commonly used are:

\begin{enumerate}
	\item Using geolocation of the device to trace its usage;
	\item Setting up the expiration period of an app;
	\item Setting up the expiration period of app pass pin;
	\item Setting up the counter of failed tries;
	\item Secured transfer politics between apps;
	\item Restricting access to the corporate app;
	\item Restricted or prohibited offline access;
	\item Logging and auditing.
\end{enumerate}

Preventing data leakage on the mobile device is a crucial security problem. Therefore, it is necessary to take additional control and protection measures for the confidential data on the mobile devices that leave the boundaries of the organization. Generally, the most sensitive and confidential data should not be permitted to be transferred to the mobile device. However, what if the SME need to allow their employees to work on such devices and even use them on the offline mode for the convenience and traffic reducing, or even for a particular characteristic of the app or the business itself?

From the theoretical point of view of this problem, there are several surveys, whose common point is the mobile cloud computing as a rapidly developing paradigm that poses many security and complexity problems \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. An analysis of the new models of mobile cloud computing and new ways of using data storage services is presented in \cite{khan2015cloud, khan2013towards}. Commonly, all the models and protection schemes concentrate on the encryption properties and either perform the computations on their own \cite{zhao2010trusted, yang2011provable} or use the cloud provider to off-load the cryptographic operations \cite{itani2010energy, ren2011lightweight}. Obviously, it is natural the mobile app cannot handle all operations securely without the assistance of a cloud provider, due to resources constraint and battery limitation.

The necessity to use schemes that function without putting load on a provider arises when it is desired to make the app less dependent on the cloud provider, i.e. corporate app continues to provide the secure access to the sensitive data without connection to the network. As discussed in \cite{khan2015cloud}, all the currently known schemes of encryption, performing the computation, either use a cloud provider \cite{zhao2010trusted}, a third party trusted agent \cite{yang2011provable} or a combination of both \cite{itani2010energy}. In some cases, they concentrate on computational complexity without taking care of user privacy and security \cite{ren2011lightweight}.
Therefore, according to \cite{khan2015cloud, tenorio2013greatest, khan2013towards}, the state-of-the-art mobile cloud security models do not consider the problem of the offline security mode.

In many cases the industrial providers of the mobile security API avoid the problem by completely forbidding the offline access to the protected app, i.e. SAP Mocona, which operates as a secure app wrapping layer \cite{Lelyveld2013}. Due to various constraints such as traffic load, travelling and ease of access, the SME business procedure may require such access. To the best of our knowledge, the offline mode security problem has not yet been deployed, neither in academia nor in the industry \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. Therefore, the main concern of this proposal is the protection of the app and its data in offline mode, when the functions of data protection cannot be offloaded to a cloud or a trusted party.

\section{The mobile security architecture}
\label{sec_the_client_security_architecture}
The approach proposed in this work describes and implements the complete lifecycle of the mobile app security infrastructure. The mobile security processes depend on the key expiry period, and are used to access the protected storage. Once the user keys expire, the user is requested to enter his valid credentials, i.e. PIN and password. The client app then sends the credentials to the server for verification. Once the new set of access keys is received, the user can access the protected files in the offline mode, without the access to the server. This means that no further communication with the server is needed until the key expires. The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:1}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=5cm]{figures/coresetoffunctionsandprotocols.eps}
	\caption{The core set of functions and protocols of the mobile cloud security infrastructure}
	\label{fig:1}
\end{figure}

@@

The cryptographic functions include the decryption (for AES and ABE) key restoration for SSS and EKE token generation procedure. While the AES-ABE hybrid encryption scheme functions both in the online and offline modes, the SSS is used only in the offline mode, while the EKE generation and executions is used in the online mode. These operations are described in Subsections \ref{sec_aes}-\ref{sec_secret} and are part of the Figure \ref{fig:1}.

The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and MOS-inspired statistics analyzer. Functions such as alerting and deleting the expired key belong to this block as well. These functions are described in Subsection \ref{sec_mos}.

The communication with server includes the separate sender and receiver to check the user credentials and receive new keys, which acts based on the EKE protocol. The remaining data like access list and files can be sent via unprotected channels.

\section{The proposed solution for offline mobile security}
\label{sec_offline_mode}
This section proposes an approach for the mobile app protection in which the security is supported both in online and offline modes. Currently and to the best of our knowledge, the systems of mobile app protection follow a model where the protected mobile app can operate only when it is connected to the cloud, which is not always convenient for the end-user. The basic principles of the mobile app protection herein proposed are:

\begin{enumerate}
	\item Optimized communication with the cloud when the app does not need to be constantly connected to the server due to the resource constraint and necessity to secure this communication;
	\item Optimized combination of the security mechanisms so that the mobile client does not need to perform complex computation like encryption and key generation due to its resource constraint;
	\item Behavioral analysis of user's operations on mobile client, which can indicate anomalous or automated activities performed by attackers.
\end{enumerate}

The most important security issues in the proposed model arise when the app goes to the offline mode and the user is still allowed to get the access to the protected SME documents. In this case, the server can neither monitor the user activity nor provide the protection methods. The security should be performed at the mobile client. Additionally, the maximum protection should be provided at the minimum resource cost.

In the online mode the mobile app uses the secure communication with the server in order to verify the validity of user’s credentials. On the contrary, the offline protection model should be approached independently. Thus, the proposal is that the authentication/authorization mechanisms in the offline mode should utilize the derived proof of the user identity. The requirements for the proof are as follows:

\begin{enumerate}
	\item The proof is derived from the previous session, in order to verify that the user is still authorized;
	\item The proof should not give access to user password, i.e. it should not be stored on the app;
	\item The proof should be temporary and have an expiration period;
	\item It should not be directly used in communication with the server, in order to prevent the malicious user from mimicking;
	\item It should be resilient to offline dictionary attacks;
	\item It has to stay effective both in the scenarios of the malicious outsider and leakage of information when the formerly authorized user leaves the group.
\end{enumerate}

Such proof cannot be stored on the mobile app, as it is not possible to guarantee its protection in the offline mode. Therefore, the most effective way to secure the proof on the offline mode without performing complex computation is to share this proof between the client and the user in a protected manner. In this case, there is no need to store the function of the user password. The additional argument against the traditional password verification is the necessity to check the PIN, which is very small, so the construction of valid one-way function resilient to the offline dictionary attacks is difficult.

The proposed architecture for offline mobile security is represented in Figure \ref{fig:5}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8.6cm]{figures/offlinesecurityarchitecture.eps}
	\caption{Proposed Architecture for Offline Mobile Security }
	\label{fig:5}
\end{figure}

@@

\begin{enumerate}
	\item \textbf{Protected Storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files. The outline of the hybrid encryption scheme is presented in the Section \ref{sec_algorithms}.
	\item \textbf{Threat Intelligence Manager (TIM)}: most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode. Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios. Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior according to the scenarios in Section \ref{sec_results}. Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones.
	\item \textbf{Key Management Center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
\end{enumerate}

@@

\section{The algorithms, key usage and data protection methods}
\label{sec_algorithms}
In the proposed approach, the kernel encryption scheme in the mobile app is a combination of several methods of security. The files are encrypted with 128/256-bit AES, while the permanent file keys are encrypted with the attribute-based encryption. The set of expiring ABE keys corresponding to the set of files accessible by user in encrypted with a single expiring AES key (KEY\_SET\_KEY). This key is expiring and is splited by the server into 4 parts (2 are stored on the device and 2 belong to the user) by the method of secure secret sharing. The encryption workflow is outlined in the Figure \ref{fig:9}.

@@

The proposed authentication system is based on the shared storing of the user key. Also, the mobile app acts as a dealer in the SSS. Using the SSS ensures that the key can only be accessed by an authenticated user. The participants of the (2, 2)-threshold SSS are the user and device. The user share $s_1(x)$ is computed based on the PIN and the PASS entered by the user. Additionally, the current time value TIME is used in the calculation of the share. Let $s(x) = d$ and $s_1(x) = f(PIN + PASS + TIME)$, where $f$ is a one-way function that transforms the data into the string of the desired length:

@@

The complexity analysis depends on the operation that the user and the mobile app perform in order to keep the app protected. In the offline mode the client app performs the following actions:

@@

The proposed concept of mobile client security has been implemented in the Storgrid protected cloud environment \cite{storgrid2016}. Therefore, the approach is correlated with the practical usability requirements:  the corporate user continues to use the mobile storage app in offline and does not need to reload the files every time the key is renewed. This methodology can be used in other mobile apps. The common advantage is that the mobile client performs the operations both in the offline and online mode and uses the key expiry and ABE to protect the privacy of the corporate data. The Table \ref{tab:01} presents the results of mobile app security workflow testing. Table \ref{tab:01} demonstrates that the mobile app decrypts the data with the acceptable speed. One can compare the speed of ABE/AES decryption (keys and actual data) and the speed of the whole decryption process.

@@

Table \ref{tab:02} demonstrates the results of testing the speed of the mobile app for receiving and displaying the list of files. The app was tested for displaying up to 1000 files. Around 100 requests were generated from 10 threads.

The mobile app receives and displays the list of files depending on the user attributes with acceptable average speed and the proposed protection scheme is usable. Table \ref{tab:03} demonstrates the results of performance testing of the keys generation procedure on server:

@@

It is possible to observe that the processing time increases with the window size decreasing, similar to the results for eigenvalue decomposition time. The longest processing time measured is lower than 200 milliseconds, even considering window size of 10 minutes or processing 6 MB of user operation log. This result represents an acceptable processing time for anomaly detection in mobile apps.

\section{Conclusion and future work }
\label{sec_conclusion}
An important security issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile apps. Once a mobile app releases the connection with the corporate cloud, no security measure implemented in the cloud infrastructure assures the protection of sensitive data stored in the mobile device. Aware of this problem and its importance, this work presented a proposal to address the offline mobile security problem combining a different cryptographic methods. Moreover, this proposed approach also prevents malicious user behavior by applying a MOS-based analytic method.
